---
title: "Solving Hansen's RBC model with lrem"
author: "Kenji Sato"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Solving Hansen's RBC model with lrem}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

The following optimization problem summarizes Gary Hansen's (1985) RBC model:

$$
\begin{aligned}
 & \max \mathbb E_0 \sum_{t=0}^{\infty}\beta^{t}\left[\log C_{t}+\chi\log\left(1-N_{t}\right)\right] \\
 & \text{subject to} \\
 & \qquad C_{t}+K_{t+1}-(1-\delta)K_{t}=z_{t}K_{t}^{\alpha}N_{t}^{1-\alpha}\\
 & \qquad\log z_{t+1}=\rho\log z_{t}+ \epsilon_{t + 1}\\
 & \qquad K_{0},z_{0}:\text{ given.} 
\end{aligned}
$$

The first-order conditions are given by ($\mathbb E_t$ omitted)

$$
\begin{aligned}
  C_{t+1}
    &=\beta C_{t}\left[1-\delta+\alpha    
      z_{t+1}K_{t+1}^{\alpha-1}N_{t+1}^{1-\alpha}\right]\\
  \chi C_{t}
    &=(1-\alpha)z_{t}K_{t}^{\alpha}N_{t}^{-\alpha}(1-N_{t})\\
  C_{t}+K_{t+1}-(1-\delta)K_{t}
    &=z_{t}K_{t}^{\alpha}N_{t}^{1-\alpha} \\
  \log z_{t+1}
    &=\rho\log z_{t} + \epsilon_{t + 1}
\end{aligned}
$$

## Steady states

The steady states are given by 

$$
\begin{aligned}
  \bar K &= \frac{(1-\alpha)q_{1}^{-\alpha}}
    {(1-\alpha)q_{1}^{1-\alpha}+\chi q_{2}}\\
  \bar C &= q_{2}\bar{K}\\
  \bar N &= q_{1}\bar{K}\\
  \bar z &= 1
\end{aligned}
$$
where 

$$
\begin{aligned}
  q_0 &= \frac{1-\beta+\beta\delta}{\alpha\beta}\\
  q_1 &= q_0^{1/(1-\alpha)}\\
  q_2 &= q_0 - \delta
\end{aligned}
$$

## Log-Linearization

By log-linearizing the first-order conditions around the steady state, 
we obtain the following linear system

$$
\begin{aligned}
  &\begin{bmatrix}\alpha(\alpha-1)q_{0} & \alpha(1-\alpha)q_{0} & -(1-\delta+\alpha q_{0})\\
  1 & 0 & 0\\
  0 & 0 & 0
  \end{bmatrix}
  \begin{bmatrix}\hat{\!K}_{t+1}\\
  \hat{\!N}_{t+1}\\
  \hat{\!C}_{t+1}
  \end{bmatrix}\\
  &\qquad 
  =\begin{bmatrix}0 & 0 & -(1-\delta+\alpha q_{0})\\
  \alpha q_{0}+1-\delta & (1-\alpha)q_{0} & -q_{2}\\
  \alpha & -\frac{\alpha+(1-\alpha)\bar{N}}{1-\bar{N}} & -1
  \end{bmatrix}\begin{bmatrix}\hat{K_{t}}\\
  \hat{N_{t}}\\
  \hat{C_{t}}
  \end{bmatrix}+\begin{bmatrix}-\alpha q_{0}\rho\\
  q_{0}\\
  1
  \end{bmatrix}\hat{z_{t}}
\end{aligned}
$$
with
$$
  \hat{\!z}_{t+1}=\rho\hat{z_{t}}+\epsilon_{t+1},
$$
where 

$$
\hat{K_{t}}:=dK_{t}/\bar{K},\ \hat{z_{t}}:=dz_{t}/\bar{z},\ 
\hat{C_{t}}:=dC_{t}/\bar{C},\ \hat{N_{t}}:=dN_{t}/\bar{N}
$$

## Parameters and coefficient matrices

We will use the following parameter values.

```{r}
alpha = 0.33
beta = 0.99
delta = 0.023
chi = 1.75
rho = 0.95
```



```{r}
q0 = (1 - beta + beta * delta) / alpha / beta
q1 = q0 ^ (1 / (1 - alpha))
q2 = q0 - delta

kbar = (1 - alpha) * q1 ^ (- alpha)
kbar = kbar / ((1 - alpha) * q0 + chi * q2)

cbar = q2 * kbar
nbar = q1 * kbar
zbar = 1
```


The matrices $E, A, B, \Phi$ are given by

```{r}
E = matrix(0, 3, 3)
A = matrix(0, 3, 3)
B = matrix(0, 3, 1)
Phi = matrix(rho, 1, 1)

E[1, 1] = alpha * (alpha - 1) * q0
E[1, 2] = alpha * q0
E[1, 3] = - (1 - delta + alpha * q0)
E[2, 1] = 1

A[1, 3] = E[1, 3]
A[2, 1] = - A[1, 3]
A[2, 2] = (1 - alpha) * q0
A[2, 3] = - q2
A[3, 1] = alpha
A[3, 2] = (- alpha - (1 - alpha) * nbar) / (1 - nbar)
A[3, 3] = -1

B[1, 1] = - alpha * q0 * rho
B[2, 1] = q0
B[3, 1] = 1
```


## Simulation using **lrem**


```{r, message=FALSE, warning=FALSE}
library("lrem")
library("ggplot2")
```

```{r}
policy <- lre_ar(A, E, B, Phi, nx = 1)

steps <- 100
out <- impulse(policy$g, policy$h, x0 = c(0, 0), t = steps, e1 = c(1, 0))
```

Since the model is log-linearized, the simulated variables represent percentage 
changes from the steady state. 

You can convert the percentage deviations into levels as follows.

```{r}
# Steady states
ss <- c(zbar, kbar, nbar, cbar)

# Convert percentage changes to levels
for (i in 1:ncol(out)) {
  out[, i] <- ss[i] * out[, i] + ss[i]
}
```

To plot these simulation result, it is convenient to convert the matrix into a
dataframe. 

```{r}
# Convert matrix to dataframe
result <- data.frame(out)
names(result) <- c("Z", "K", "N", "C")
result["t"] <- 0:(steps - 1)

# Plotting
ggplot(result) + geom_line(aes(x = t, y = K))
ggplot(result) + geom_line(aes(x = t, y = Z))
ggplot(result) + geom_line(aes(x = t, y = N))
ggplot(result) + geom_line(aes(x = t, y = C))

```
