---
title: "Final Assignment"
author: "Kenji Sato"
date: '`r format(Sys.time(), "%d %B, %Y")`'
output:
  bookdown::html_document2: default
  bookdown::pdf_document2:
    fig_caption: yes
    includes:
      in_header: asset/latex/custom.sty
    keep_tex: no
    latex_engine: pdflatex
    number_sections: yes
    template: null
    toc: no
documentclass: scrartcl
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Deadline and submission

June 9, 2017, 18:00  
Late submission not accepted.

# Assignment

Complete your **lrem** project that you launched for `hw08` and do a simulation in vignettes. 

In addition, do the following.

* Write `README.md`. The content must include 
    * installation manual 
      (`devtools::use_readme_rmd()` will help); and
    * a short description about what the package does.
* Write a short paper accompanying the package (as a _vignette_).

## Specifications

Your package should be able to solve a linear rational expectations (LRE) models. 

### A. Autonomous LRE System {-}

Your package solves LRE model without inputs.

\begin{equation}
E \mathbb E_t 
\begin{bmatrix}
  x_{t+1}^1 \\
  x_{t+1}^2
\end{bmatrix}
=
A
\begin{bmatrix}
  x_{t}^1 \\
  x_{t}^2
\end{bmatrix}
(\#eq:lre-auto)
\end{equation}

where $x^1 \in \mathbb R^{n_1}$ is the vector of predetermined variables and $x^2 \in \mathbb R^{n_2}$ the vector of non-predetermined variables.

The solution to Eq. \@ref(eq:lre-auto)
\begin{equation}
  \begin{aligned}
    x_{t+1}^1 &= h(x_t^1) + \xi_{t+1}\\
    x_{t}^2 &= g(x_t^1)
  \end{aligned}
  (\#eq:lre-sol)
\end{equation}
is characterized by two functions $(g, h)$. See the lecture slides for the formula.

### Suggested API {-}

We would suggest that you implement this algorithm with the following API. 
Whether or not you use this API, document usage properly. 

-----------------------------------------------------------------
`lre_auto`      
------------- ---------------------------------------------------
Purpose       Compute $(g, h)$ from $(A, E, n_1)$
  
Arguments     `A`: matrix $A$\
              `E`: matrix $E$\
              `nx`: number of predetermined variables, $n_1$, OR\
              `x0`: initial vector, $x_0^1$
               
Return Value  `list(g = g, h = h)`: list of two functions 
              $(g, h)$ in \@ref(eq:lre-sol)
------------------------------------------------------------------

Users should be able to use this function with the following code 
combination after properly defined `A` and `E`.

```{r, eval = FALSE}
sol <- lre_auto(A, E, nx = 3)
g <- sol$g
h <- sol$h
```


### B. LRE System with AR Inputs {-}

Let $\Phi$ be a (strictly) stable matrix and $\epsilon$ be i.i.d. shocks.

\begin{equation}
\begin{aligned}
E \mathbb E_t 
\begin{bmatrix}
  x_{t+1}^1 \\
  x_{t+1}^2
\end{bmatrix}
&=
A
\begin{bmatrix}
  x_{t}^1 \\
  x_{t}^2
\end{bmatrix}
+
B u_t\\
u_{t+1} 
&= \Phi u_t + \epsilon_{t+1}\\
\end{aligned}
(\#eq:lre-ar)
\end{equation}

This system satisfies 

\begin{equation}
  \begin{bmatrix}
    I & O \\
    O & E
  \end{bmatrix}
  \mathbb E_t 
  \begin{bmatrix}
    u_{t+1}\\
    \hline
    x_{t+1}^1 \\
    x_{t+1}^2
  \end{bmatrix}
  =
  \begin{bmatrix}
    \Phi & O \\
    B    & A
  \end{bmatrix}
  \begin{bmatrix}
    u_{t}\\
    \hline
    x_{t}^1 \\
    x_{t}^2
  \end{bmatrix}
  (\#eq:lre-ar2)
\end{equation}

and under the assumptions discussed in the class \@ref(eq:lre-ar2) is solved by 
\begin{equation}
  \begin{aligned}
    \begin{bmatrix}
    u_{t+1} \\
    x_{t+1}^1 
    \end{bmatrix}
    &=
    h(u_t, x_t^1)
    +
    \begin{bmatrix}
    I \\
    0 
    \end{bmatrix}
    \epsilon_{t+1}
    \\
    x_t^2 &= g(u_t, x_t^1),
  \end{aligned}
  (\#eq:lre-ar-sol)
\end{equation}

where $g$ and $h$ here are computed by the same algorith used for autonomous system but with the extended coefficient matrices.

(If you feel unhappy with the order of $(u_t, x_t^1)$ and want it to be 
$(x_t^1, u_t)$, you may arrange it freely. You will need to define 
the coefficient matrices more carefully.)

Provide an R function that computes $(g, h)$ in \@ref(eq:lre-ar-sol).

### Suggested API {-}

We would suggest that you implement this algorithm with the following API. 
Whether or not you use this API, document usage properly. 

-----------------------------------------------------------------
`lre_ar`      
------------- ---------------------------------------------------
Purpose       Compute $(g, h)$ from $(A, E, B, \Phi, n_1)$
  
Arguments     `A`: matrix $A$\
              `E`: matrix $E$\
              `B`: matrix $B$\
              `Phi`: matrix $\Phi$\
              `nx`: number of predetermined variables, $n_1$, OR\
              `x0`: initial vector, $x_0^1$
               
Return Value  `list(g = g, h = h)`: list of two functions 
              $(g, h)$ in \@ref(eq:lre-ar-sol)
------------------------------------------------------------------


Hint: Let $n_\epsilon$ be the size of exogenous shocks. You can construct
extended coefficient matrices in \@ref(eq:lre-ar2) from `A`, `E`, `B`
and `Phi`. Pass the obtained matrices to `lre_auto`.


### C. Simulate {-}

You need to extend that old `simulate` function to accept shock 
process $\xi_t$ or $\epsilon_t$.

### Suggested API {-}

We would suggest that you implement simulation function with the 
following API. Whether or not you use this API, document usage properly. 

-----------------------------------------------------------------
`simulate`      
------------- ---------------------------------------------------
Purpose       Simulation given $(g, h, x_0^1, \xi)$
  
Arguments     `g`: function $g$ in \@ref(eq:lre-sol) or 
                   \@ref(eq:lre-ar-sol)\
              `h`: function $h$ in \@ref(eq:lre-sol) or 
                   \@ref(eq:lre-ar-sol)\
              `x0`: initial $x_0^1$; understood as 
                    $(u_0, x_0^1)$ in case of \@ref(eq:lre-ar-sol)\
              `t`: Integer, simulation length\
              `e`: vector or matrix, each row `e[k, ]` corresponds
                   to $\xi_{k+1}$.
               
Return Value  `out`: matrix of simulation output 
------------------------------------------------------------------

Either one of `e` or `t` can accept `NULL`. 

* If `e` is `NULL`, it is understood as zero shocks.
* If `t` is `NULL`, `nrow(e) + 1` might be used as `t`.

If `nrow(e)` is smaller than `t - 1`, you might need to extend matrix `e` to have `t - 1` rows and fill the empty elements with zeros.

### D. Impulse Response {-}

Let $x_0^1$ and $x_0^2$ be on the steady state, i.e., they are zero. 

To make code as simple as possible, let's define an impulse response 
as a response to 

\begin{equation}
  \xi_1 \neq 0, \xi_2 =0, \dots
\end{equation}

Users are responsible to set, for example,

\begin{equation}
  \xi_1 = \begin{bmatrix}
    0 \\ 1\\ 0
  \end{bmatrix}
\end{equation}

### Suggested API {-}

We would suggest that you implement this algorithm with the following API. 
Whether or not you use this API, document usage properly. 

-----------------------------------------------------------------
`impulse`      
------------- ---------------------------------------------------
Purpose       Impulse response given $(g, h, \xi_1)$
  
Arguments     `g`: function $g$ in \@ref(eq:lre-sol) or 
                   \@ref(eq:lre-ar-sol)\
              `h`: function $h$ in \@ref(eq:lre-sol) or 
                   \@ref(eq:lre-ar-sol)\
              `x0`: steady state. For linear models, it's the 
                    origin.\
              `t`: Integer, simulation length\
              `e1`: i-th element `e[i]` corresponds to a shock 
                   to i-th endogenous variable
               
Return Value  `out`: matrix of simulation output 
------------------------------------------------------------------

It seems redundant to pass `x0` because it should be zero.
This is due to the unfortunate fact that our `g` and `h` don't 
have information on the system size.


## Exmaple 1

See [lrem-hansen-rbc.html](lrem-hansen-rbc.html) 

Try to replicate this example.


## Example 2

Simulate the canonical New-Keynesian model that consists of the following three
equations. 

**Phillips curve (AS curve)**
$$\pi_{t}	=\beta\mathbb{E}_{t}\pi_{t+1}+\kappa x_{t}+u_{t}^{S}$$

**IS curve**
$$x_{t}=\mathbb{E}_{t}x_{t+1}-\frac{1}{\sigma}\left(i_{t}-\mathbb{E}_{t}\pi_{t+1}\right)+u_{t}^{D}$$


**Monetary policy rule**
$$
i_{t}=\alpha\pi_{t}+\iota x_{t}
$$


Assume AR shocks with $|\rho_{S}|$, $|\rho_{D}|<1$, $\mathbb{E}_{t}\epsilon_{t+1}^{S}=0$, and $\mathbb{E}_{t}\epsilon_{t+1}^{D}=0$: 
$$
\begin{aligned}
  u_{t+1}^{S}	&=\rho_{S}u_{t}^{S}+\epsilon_{t+1}^{S}\\
  u_{t+1}^{D}	&=\rho_{D}u_{t}^{D}+\epsilon_{t+1}^{D}
\end{aligned}
$$

* $\pi_{t}$: log inflation rate
* $x_{t}$: output gap
* $i_{t}$: log gross nominal interest rate
* $\sigma$: intertemporal elasticity of substitution
* $\beta$: discount rate
* $\kappa$: slope of the Phillips curve
* $\alpha$ and $\iota$: monetary policy parameters

Simulate the impulse response with the following parameter values.

```{r}
alpha <- 2
iota <- 0
beta <- 0.99
sigma <- 1
kappa <- 0.132
rhoS <- 0.9
rhoD <- 0.9
```


All three endogenous variables, $\pi$, $x$ and $i$, are non-predetermined.


# Assessment

The TA and lecturer (henceforth, the users) will ask themselves the following questions.

* **Installation**: Does installation go smoothly?
* **Vignettes**: Is the vignette(s) written reasonably well?
* **Functionality**: Does the package work as expected?
* **Documentation**: Is the package well documented?
* **Format**: Does the code written (aesthetically) well?

## Installation

The users will install your package 
on their computers following the instruction provided in `README.md`.

`README.md` should provide a sunccinct description about who wants to 
use the package and what it provides to solve a problem. 
Desirably, it provides a short example. Long example should be written 
in a vignette.

## Vignettes

In the vignette(s), you will discuss how the package is supposed to be 
used to solve a real problem. 

The users check what vignettes the package has by 

```r
vignette(package = "lrem")
```

Then they read them with

```r
vignette("lrem-vignette-topic", package = "lrem")
```

Since your package contains several functions, the users want to know 
how those functions can be combined to solve their tasks. 

See <http://r-pkgs.had.co.nz/vignettes.html> for more details.


## Functionality

In general, the package should be able to solve the problems 
`README.md` and vignettes declare that it does.

The users will

* redo the computation described in the vignettes; and 
* try to solve problems they have, which are unknown to the package 
  creator.

Since this project is for a course, your package is supposed to 
have the functionalities discussed in Section \@ref(assignment).

## Documentation

The users read the help page (for example, by `?lre_auto`) very
very often. Documentation is important. See <http://r-pkgs.had.co.nz/man.html>
to learn how to write documentation.


## Format

Sensible users read the code (at least partially) to become confident 
that the package provides what they really want. If your code doesn't 
stick to a consistent and reasonable style, the users might stop 
using it because they think the package might be unreliable. 

Format is important also because ill-formatted code is difficult to maintain.

Follow this style <http://adv-r.had.co.nz/Style.html> if you have no particular style yet.

# Remarks 

## Collaboration

Because the repositories you make are public, everybody else can see your work. I don't deny you to learn from other students' works but keep in mind that their works are copyright protected unless otherwise explicitly stated.

Be a sensible user not a pirate.

## Need help?

If you have questions about this assignment, ask on Slack. 

